---
title: "Assg_1"
author: "Sonali Kalthur"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

##Load required library packages
```{r packages}
library(cluster)
library(ggplot2)
library(ggfortify)
library(caret)
library(factoextra)
```

##Reading the survey data
```{r}
setwd("~/Documents/Semester2/Computationalstatistics/tutorial/Survey_data")
survey_list <- read.csv("survey_final_clean_transposed_v1.1.csv", header = T, na.strings=c(""," ","NA", "-"))

# Column Names changed to abbreviated form
names(survey_list) <- c("ID", "professional_experience", "job_satisfaction", "professional_background", "stats_experience", "programming_experience", "amount_entertainment", "amount_study", "enough_effort", "amount_funreading", "country_born", "month_born", "income", "statisticians_lie", "visualisation_factors", "visualisation_vs_datacleaning", "UOS_satisfaction", "salary_wish", "favourite_colour", "street_number", "highschool_satisfaction", "eye_colour", "pokemons_caught", "work_experience", "undergrad_major", "current_degree", "stats_level", "statistician", "professional_parents", "height", "age", "target_industry", "countries_visited", "countries_livedin", "parental_language", "expected_grade", "favourite_movie", "like_chocolate", "tennis_vs_rugby", "STAT5002_guess", "R_guess", "IT_guess", "maths_guess", "trump_threat", "earliest_memory", "countries_totravel", "target_industry2", "home_value", "home_satisfaction", "mostfamiliar_programming", "age_independent", "favourite_number")
```

##Analyzing the times spent in a week by participants to experience level of statistics
```{r}
##Outlier treatment for How many days do you spend on entertainment each week?
for(i in seq(survey_list$amount_entertainment)) { 
  if(survey_list$amount_entertainment[i] > 7) { 
    survey_list$amount_entertainment[i] = 7
  }}

##Converting the hours in the column 9 to days
for(i in seq(survey_list$amount_funreading)) {
    survey_list$amount_funreading[i] = survey_list$amount_funreading[i]/24
  }


##Creating a subset and replacing na with 0
subset_2<- survey_list[,c('amount_entertainment','amount_study','amount_funreading')]
subset_2[is.na(subset_2)]<- 0

##Checking for optimal K value
Koptimal_val <- (length(subset_2)-1)*sum(apply(subset_2,2,var))
for (i in 1:20) Koptimal_val[i] <- sum(kmeans(subset_2,centers=i)$withinss)
plot(1:20, Koptimal_val, type="b", xlab="Number of Clusters",
  ylab="Dataset Value")

##applying Kmeans Clustering on the data
subset_cluster<- kmeans(subset_2,4,nstart=1)
table(subset_cluster$cluster,survey_list$expected_grade) 

##Plotting the Cluster
#install.packages('ggfortify')
#library(ggfortify)
autoplot(subset_cluster,data=subset_2)

##Validating the cluster using silhoutte width 
sil <- silhouette(subset_cluster$cluster, dist(subset_2))
head(sil[, 1:3], 10)
plot(sil, main ="Silhouette plot - K-means")
fviz_silhouette(sil)

# Summary of silhouette analysis
sil.summary <- summary(sil)
# Average silhouette width of each cluster
sil.summary$clus.avg.widths
# The total average (mean of all individual silhouette widths)
sil.summary$avg.width
# The size of each clusters
sil.summary$clus.sizes
```


## Clustering people with similar experience(background) to check thier salary exceptation
```{r}
#Creating subset
subset_3<- survey_list[,c('stats_experience','programming_experience','professional_experience')]

##Treating missing values for Income and replacing stats_experience 0 with NaN

#subset_3$income[subset_3$income==0]<-NaN
subset_3$stats_experience[is.na(subset_3$stats_experience) ] <- 0
#subset_3$income <- with(subset_3,ave(subset_3$income,subset_3$professional_experience,FUN = function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))))
#subset_3$income[is.nan(subset_3$income)] <- 0

##Doing a scale
max_val <-apply(subset_3,2,max)
min_val <-apply(subset_3,2,min)
subset_3_scale<- scale(subset_3, center=min_val, scale = max_val-min_val)
```

##Kmeans Clustering
```{r}
subset_cluster1<- kmeans(subset_3_scale,5,nstart=1)
subset_cluster1$cluster
#plot(subset_3_scale,col=subset_cluster1$cluster)
#points(subset_cluster1$centers, pch=16)
autoplot(subset_cluster1,data=subset_3_scale)
table(subset_cluster1$cluster,survey_list$salary_wish) 

##Checking for optimal K value
Koptimal_val <- (length(subset_3)-1)*sum(apply(subset_3,2,var))
for (i in 1:20) Koptimal_val[i] <- sum(kmeans(subset_3,centers=i)$withinss)
plot(1:20, Koptimal_val, type="b", xlab="Number of Clusters",
  ylab="Dataset Value")

##Validating the cluster using silhoutte width (the one clustered has no group because its a blank)
sil <- silhouette(subset_cluster1$cluster, dist(subset_3_scale))
head(sil[, 1:3], 10)
plot(sil, main ="Silhouette plot - K-means")
fviz_silhouette(sil)

# Summary of silhouette analysis
sil.summary <- summary(sil)
# Average silhouette width of each cluster
sil.summary$clus.avg.widths
# The total average (mean of all individual silhouette widths)
sil.summary$avg.width
# The size of each clusters
sil.summary$clus.sizes
```


##People with different professional backgrounds are skilled in both programming and statistics

## Clustering people with similar experience(background) to check thier salary exceptation
```{r}
#Creating subset
subset_4<- survey_list[,c('stats_experience','programming_experience','professional_experience')]


##replacing stats_experience 0 with NaN
subset_4$stats_experience[is.na(subset_4$stats_experience) ] <- 0

##Converting total number of professional years year to days
subset_4$professional_experience<- subset_4$professional_experience*365 

##Doing a scale
#max_val <-apply(subset_3,2,max)
#min_val <-apply(subset_3,2,min)
#subset_3_scale<- scale(subset_3, center=min_val, scale = max_val-min_val)
```

##Kmeans Clustering
```{r}
subset_cluster2<- kmeans(subset_4,5,nstart=1)
subset_cluster2$cluster
#plot(subset_3_scale,col=subset_cluster1$cluster)
#points(subset_cluster1$centers, pch=16)
autoplot(subset_cluster2,data=subset_4)
table(subset_cluster2$cluster,survey_list$professional_background) 

##Checking for optimal K value
Koptimal_val <- (length(subset_4)-1)*sum(apply(subset_4,2,var))
for (i in 1:20) Koptimal_val[i] <- sum(kmeans(subset_4,centers=i)$withinss)
plot(1:20, Koptimal_val, type="b", xlab="Number of Clusters",
  ylab="Dataset Value")

##Validating the cluster using silhoutte width (the one clustered has no group because its a blank)
sil <- silhouette(subset_cluster2$cluster, dist(subset_4))
head(sil[, 1:3], 10)
plot(sil, main ="Silhouette plot - K-means")
fviz_silhouette(sil)

# Summary of silhouette analysis
sil.summary <- summary(sil)
# Average silhouette width of each cluster
sil.summary$clus.avg.widths
# The total average (mean of all individual silhouette widths)
sil.summary$avg.width
# The size of each clusters
sil.summary$clus.sizes
```

##People with high salary are they satisfied with their current place of living?


## Clustering people with similar experience(background) to check thier salary exceptation
```{r}
#Creating subset
subset_4<- survey_list[,c('programming_experience','professional_experience','job_satisfaction')]

##replacing stats_experience 0 with NaN
subset_4$stats_experience[is.na(subset_4$stats_experience) ] <- 0


##Doing a scale
#max_val <-apply(subset_3,2,max)
#min_val <-apply(subset_3,2,min)
#subset_3_scale<- scale(subset_3, center=min_val, scale = max_val-min_val)
```

##Kmeans Clustering
```{r}
subset_cluster2<- kmeans(subset_4,4,nstart=1)
subset_cluster2$cluster
#plot(subset_3_scale,col=subset_cluster1$cluster)
#points(subset_cluster1$centers, pch=16)
autoplot(subset_cluster2,data=subset_4)
table(subset_cluster2$cluster,survey_list$professional_background)


##Checking for optimal K value
Koptimal_val <- (length(subset_4)-1)*sum(apply(subset_4,2,var))
for (i in 1:20) Koptimal_val[i] <- sum(kmeans(subset_4,centers=i)$withinss)
plot(1:20, Koptimal_val, type="b", xlab="Number of Clusters",
  ylab="Dataset Value")

##Validating the cluster using silhoutte width (the one clustered has no group because its a blank)
sil <- silhouette(subset_cluster2$cluster, dist(subset_4))
head(sil[, 1:3], 10)
plot(sil, main ="Silhouette plot - K-means")
fviz_silhouette(sil)

# Summary of silhouette analysis
sil.summary <- summary(sil)
# Average silhouette width of each cluster
sil.summary$clus.avg.widths
# The total average (mean of all individual silhouette widths)
sil.summary$avg.width
# The size of each clusters
sil.summary$clus.sizes
```


##Clustering all the rating scale value to check how it differs from each country of origin
```{r}
#Creating subset
subset_5<- survey_list[,c('job_satisfaction','UOS_satisfaction','highschool_satisfaction','home_satisfaction')]

##replacing stats_experience 0 with NaN
subset_5$job_satisfaction[is.na(subset_5$job_satisfaction) ] <- 0
subset_5$UOS_satisfaction[is.na(subset_5$UOS_satisfaction) ] <- 0
subset_5$highschool_satisfaction[is.na(subset_5$highschool_satisfaction) ] <- 0
#subset_5$trump_threat[is.na(subset_5$trump_threat) ] <- 0
subset_5$home_satisfaction[is.na(subset_5$home_satisfaction) ] <- 0




##Doing a scale
max_val <-apply(subset_5,2,max)
min_val <-apply(subset_5,2,min)
subset_3_scale<- scale(subset_5, center=min_val, scale = max_val-min_val)
```



```{r}
subset_cluster3<- kmeans(subset_3_scale,5,nstart=1)
subset_cluster3$cluster
#plot(subset_3_scale,col=subset_cluster1$cluster)
#points(subset_cluster1$centers, pch=16)
autoplot(subset_cluster3,data=subset_3_scale)
table(subset_cluster3$cluster,survey_list$country_born) 

##Checking for optimal K value
Koptimal_val <- (length(subset_3_scale)-1)*sum(apply(subset_3_scale,2,var))
for (i in 1:20) Koptimal_val[i] <- sum(kmeans(subset_3_scale,centers=i)$withinss)
plot(1:20, Koptimal_val, type="b", xlab="Number of Clusters",
  ylab="Dataset Value")

##Validating the cluster using silhoutte width (the one clustered has no group because its a blank)
#sil <- silhouette(subset_cluster3$cluster, dist(subset_3_scale))
sil <- silhouette(subset_cluster3$cluster, dist(subset_3_scale))
head(sil[, 1:3], 10)
plot(sil, main ="Silhouette plot - K-means")
fviz_silhouette(sil)

# Summary of silhouette analysis
sil.summary <- summary(sil)
# Average silhouette width of each cluster
sil.summary$clus.avg.widths
# The total average (mean of all individual silhouette widths)
sil.summary$avg.width
# The size of each clusters
sil.summary$clus.sizes
```

##Clustering the experience in statistics and currently using statistics in work to check with the experience with statistics
```{r}
#Creating subset
subset_6<- survey_list[,c('amount_entertainment','amount_funreading','height','countries_livedin','age_independent')]

subset_6$amount_funreading[is.na(subset_6$amount_funreading) ] <- 0
subset_6$age_independent[is.na(subset_6$age_independent) ] <- 0
subset_6$height[is.na(subset_6$height) ] <- 0
#Conversting the dtatistician value from Yes and No to 1 and 0
#subset_6$eye_colour <- unfactor(subset_6$eye_colour , levels=c("Brown","Blue","Black","Green","Hazel","Dark Brown"),
#labels = c(1, 2, 3, 4, 4, 1))


#subset_6$statistician <- ifelse(subset_6$statistician=='Y',1,0)
#subset_6$statistician[is.na(subset_6$statistician) ] <- 0

##Doing a scale
max_val <-apply(subset_6,2,max)
min_val <-apply(subset_6,2,min)
subset_6_scale<- scale(subset_6, center=min_val, scale = max_val-min_val)
```



##Kmeans Clustering
```{r}
subset_cluster4<- kmeans(subset_6_scale,3,nstart=1)
subset_cluster4$cluster
#plot(subset_3_scale,col=subset_cluster1$cluster)
#points(subset_cluster1$centers, pch=16)
autoplot(subset_cluster4,data=subset_6_scale)
table(subset_cluster4$cluster,survey_list$country_born) 

##Checking for optimal K value
Koptimal_val <- (length(subset_6_scale)-1)*sum(apply(subset_6_scale,2,var))
for (i in 1:20) Koptimal_val[i] <- sum(kmeans(subset_6_scale,centers=i)$withinss)
plot(1:20, Koptimal_val, type="b", xlab="Number of Clusters",
  ylab="Dataset Value")

##Validating the cluster using silhoutte width (the one clustered has no group because its a blank)
#sil <- silhouette(subset_cluster3$cluster, dist(subset_3_scale))
sil <- silhouette(subset_cluster4$cluster, dist(subset_6_scale))
head(sil[, 1:3], 10)
plot(sil, main ="Silhouette plot - K-means")
fviz_silhouette(sil)

# Summary of silhouette analysis
sil.summary <- summary(sil)
# Average silhouette width of each cluster
sil.summary$clus.avg.widths
# The total average (mean of all individual silhouette widths)
sil.summary$avg.width
# The size of each clusters
sil.summary$clus.sizes
```
##Density estimation for Income and Height 
##Smoothing by finding the alpha value
```{r}

ggplot(survey_list, aes(x=income)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, col="")

ggplot(survey_list, aes(x=height,y=..density..)) + 
 geom_histogram(colour="black", fill="grey",bin=20) +
 geom_density(aes(col="6.3"),bw=6.3,alpha=.2) +
 geom_density(aes(col="2"),bw=2.0,alpha=.2) +
 xlab("Height")+
 ylab("") 
 scale_color_manual(name="bandwidth", values=c("red","blue"))
```